### 2022-09-08 _dev/supported_models.Rmd
Roxygen updates

---
title: "Model Support"
author: "Noah Greifer"
date: "2022-08-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Supported models

* `stats::lm()`

For `mlm` models (multiple outcomes), coefficients are named `y1:x1` and `y2:x1` as they are in the covariance matrix when supplied through `coefs` argument of `FUN` in `sim_apply()`; when extracted from `fit`, they are in a matrix as usual.

* `stats::glm()`

* `MASS::glm.nb()`

* `survey::svyglm()`

* `estimatr::lm_robust()`

See `lm()` about multiple outcomes.

* `fixest::feols()`, `fixest::feglm()`, `fixest::femlm()`, etc.

Cannot accept `fixest.multi` models, i.e., resulting from multiple models being fit in the same `feols()` call. In IV, only second stage coefficients and covariance are extracted.

* `betareg::betareg()`

* `logistf` models (`logistf::logistf()`, `logistf::flic()`, `logistf::flac()`)

* `geepack::geeglm()`

* `rms::ols()`, `rms::lrm()`

* Multiply imputed data (fit list or `mids`)

Need to add tests and examples

## Models to add

* `survival` models (incl. `rms::cph()`)

* `ivreg` models

* `gam` models

* `estimatr::iv_robust()`

Only operates on second stage model. NOTE: bug in `estimatr` preventing full use.

### Way down the road

* Mixed effects models (`nlme`, `lme4`, `glmmTMB`)

* Multinomial regression (`nnet::multinom()`, `mlogit::mlogit()`, `mclogit::mblogit()`)

* Ordinal regression (`MASS::polr()`, `rms::olr()`, `geepack::ordgee()`)

* `brms::brms()`

Probably want an `as.sim()` method to use posterior as simulated values. May also want to trigger a Bayesian pipeline.

* `ordbetareg::ordbetareg()`

Returns a `brms` object.

## Adding new models

* Add `transform_coefs()` (probably okay unless `coef()` doesn't return a vector)
* Add `get_vcov()` (probably okay)
* Add `get_df()` (probably set to `Inf` unless model is a linear model fit with OLS)
* Add `get_coef_template()` (probably okay unless coefficients stored in unusual way)
* Add `untransform_coefs()` (probably okay unless coefficients stored in unusual way)
* Add `get_coef_location()` (probably `"coefficients"`)
* Add `simbased_predict()` (probably okay unless multivariate)
* Add test to `test-sim.R` and `test-sim_ame.R`